---
title: "Relapse duration"
output: html_notebook
---

Investigate what ODE structure would be required for a delayed relapse.

```{r setup}
library(tidyverse)
library(readxl)
library(patchwork)
```

Investigate Guojing's relapse data

```{r}
relapse = read_excel("../data/china/Guantang Relapse Data - GJY.xlsx", "Relapse expanded", skip=1) %>%
  mutate_all(as.character) %>%
  filter(!str_detect(Volunteer, "participants")) %>%
  pivot_longer(cols=matches("Incub|relapse"), names_to="event", values_to="delay") %>%
  mutate(delay = delay %>% as.numeric(),
         event = ifelse(str_detect(event, "Inc"), "0 Incubation period", event)) %>%
  drop_na(delay)

ggplot(relapse, aes(x=delay, fill=event)) +
  geom_histogram(binwidth=30) +
  facet_wrap(vars(event), ncol=1)
```

Individual level

```{r}
indv = relapse %>%
  arrange(Volunteer, event) %>%
  mutate(event_num = as.numeric(substr(event, 1, 1)),
         event = fct_relevel(event, sort(levels(event)))) %>%
  group_by(Volunteer) %>%
  mutate(marginal_delay = delay - lag(delay, default=0)) %>%
  select(Volunteer, event, event_num, delay, marginal_delay) %>%
  ungroup() %>%
  mutate(is_long = event_num <= 1 & marginal_delay >= 120,
         is_short = event_num > 1 | event_num == 1 & marginal_delay < 120,
         event_type = case_when(is_long ~ "long",
                                is_short ~ "short",
                                T ~ "primary") %>%
           fct_relevel(c("long", "short", "primary")))

p1 = ggplot(indv, aes(x=delay/365.25*12, y=event, group=Volunteer)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(breaks = 1:15) +
  labs(subtitle = "Individual volunteer events after primaryation")

p2 = ggplot(indv, aes(x=marginal_delay/365.25*12, y=event, color=event_type, group=Volunteer)) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(breaks = 1:15) +
  labs(subtitle = "Individual volunteer delays after previous event")

p1 / p2 + plot_layout(guides = "collect")
```

Can we use the above graph to calibrate?

```{r}
short_relapses = indv %>%
  filter(is_short)
long_relapses = indv %>%
  filter(is_long)

stats = indv %>%
  group_by(event_type) %>%
  summarise(mean_delay = mean(marginal_delay),
            var = var(marginal_delay))

# Therefore
f = 1/stats$mean_delay[2]
delta = 1/(stats$mean_delay[1] - stats$mean_delay[2]) # delta = rate/shape
var_gamma = stats$var[1] - stats$var[2]
# mean = k*theta
# var = mean*theta
# delta = theta/k
# theta = mean/var
# k = theta/delta = mean/var/delta
# k = (stats$mean_delay[1] - stats$mean_delay[2]) / var_gamma / delta

# Note: based on this, 1/f = 64.2, 1/delta = 158.1, n_compartments = 24.6
```

Algebra
m = a/b
v = a/b^2
v = m/b
b = mean/variance
a = m*b
a = mean^2/variance

```{r}
sample_mean = stats$mean_delay[1] - stats$mean_delay[2]
sample_var = var_gamma
# shape = 50 # number of compartments
shape = sample_mean^2 / sample_var
# rate = 1/2 # rate = 1 / mean time spent in each compartment
rate = sample_mean / sample_var


time = seq(0, 365.25*2, length.out=1000)
tibble(time=time,
       density = dgamma(time, round(shape), 1/(round(1/rate)))) %>%
  ggplot(aes(x=time/365.25*12, y=density)) +
  geom_line()

# n = 1000
# samples = tibble(t1 = rgamma(n, shape, rate),
#                  t2 = rexp(n, 1/10),
#                  t = t1 + t2)
# ggplot(samples, aes(x=t)) +
#   geom_density()
```

```{r}
message("1/f: ", 1/f)
message("1/delta: ", 1/delta)
message("shape: ", shape)
```

Erlang dormancy with an exponential final relapse period

```{r}
# library(bayesmeta)
# total_dist = convolve(function(x) {
#   dgamma(x, shape=dist_erlang$k, rate=dist_erlang$rate)
# },
# function(x) {
#   dexp(x, rate=dist_exp$lambda)
# },
# # delta = 0.000001,
# # epsilon = 0.0001
# )
analytic = tibble(
  tt = seq(400),
  Dormancy = dgamma(tt, shape=dist_erlang$k, rate=dist_erlang$rate),
  Latency = dexp(tt, rate=dist_exp$lambda),
  `Dormancy + latency` = total_dist$density(tt)
) %>%
  pivot_longer(-tt) %>%
  mutate(name = fct_inorder(name))

sample_relapses = function(delta, n_compartments, f) {
  dist_erlang = list(k = n_compartments, mean = 1/delta)
  dist_erlang$rate = with(dist_erlang, k / mean)
  dist_exp = list(lambda=f, mean = 1/f, var=1/f^2)
  
  n = 100000
  samples = tibble(i = seq_len(n),
                   Dormancy = rgamma(n, shape=dist_erlang$k, rate=dist_erlang$rate),
                   Latency = rexp(n, rate=dist_exp$lambda),
                   `Dormancy + latency` = Dormancy + Latency) %>%
    pivot_longer(-i) %>%
    mutate(name = fct_inorder(name))
  
  samples %>%
    group_by(name) %>%
    summarise(mean = mean(value),
              sd = sd(value))
  
  quantile(samples$value[samples$name == "Dormancy + latency"], c(0.25, 0.75))
  mean(samples$value[samples$name == "Dormancy + latency"])
  sd(samples$value[samples$name == "Dormancy + latency"])
  return(samples)
}

samples_list = list(
  `δ=1/162, k=11` = sample_relapses(1/162, 11, 1/72),
  `δ=1/240, k=22` = sample_relapses(1/240, 22, 1/72),
  `δ=1/240, k=100` = sample_relapses(1/240, 100, 1/72),
  `δ=1/240, k=1000` = sample_relapses(1/240, 1000, 1/72),
  `TODO` = sample_relapses(delta, shape, f))
names(samples_list)[5] = paste0("δ=1/", round(1/delta), ", k=", round(shape), ", f=1/", round(1/f))
samples = samples_list %>%
  bind_rows(.id = "Scenario") %>%
  mutate(Scenario = fct_inorder(Scenario))

relapse_plot = samples %>%
  ggplot(aes(x = value/365.25*12, fill = name)) +
  geom_density(aes(y=..scaled..), alpha=0.8, color="transparent", adjust=0.01) +
  # facet_wrap(vars(name), ncol=1) +
  facet_grid(rows = vars(Scenario), cols = vars(name)) +
  coord_cartesian(xlim=c(0, 16)) +
  scale_x_continuous(breaks = c(0:18)) +
  theme(legend.position = "none") +
  labs(x = "Months", y = NULL)
relapse_plot

# analytic %>%
#   ggplot(aes(x = tt/365.25*12, y = value, fill = name)) +
#   geom_area(alpha=0.8, color="transparent") +
#   facet_wrap(vars(name), ncol=1) +
#   coord_cartesian(xlim=c(0, 12)) +
#   scale_x_continuous(breaks = c(0:12)) +
#   labs(x = "Months", y = "Probability density (analytic)") +
#   theme(legend.position = "none")

# ggsave("../plots/analytic_relapse_duration_delta220.png", width=8, height=4)
```

Plot original data with fit distribution

```{r}
china_fit = samples %>%
  filter(Scenario == paste0("δ=1/", round(1/delta), ", k=", round(shape), ", f=1/", round(1/f))) %>%
  mutate(event_type = case_match(name,
                                 "Latency" ~ "short",
                                 "Dormancy + latency" ~ "long") %>%
           fct_relevel(levels(indv$event_type))) %>%
  right_join(indv, by="event_type") %>%
  mutate(event_type = case_match(event_type,
                                 "short" ~ "Latency (short delay)",
                                 "long" ~ "Dormancy + latency\n(long delay)",
                                 "primary" ~ "No delay") %>%
           fct_inorder() %>%
           fct_relevel("No delay"),
         event = event %>% str_remove("^0 ") %>%
           fct_relevel("Incubation period"))

indv_2 = indv %>%
  mutate(event_type = case_match(event_type,
                                 "short" ~ "Latency (short delay)",
                                 "long" ~ "Dormancy + latency\n(long delay)",
                                 "primary" ~ "No delay") %>%
           fct_inorder() %>%
           fct_relevel("No delay"),
         event = event %>% str_remove("^0 ") %>%
           fct_relevel("Incubation period"))

ggplot() +
  # geom_point(alpha = 0.5) +
  ggridges::geom_density_ridges(data=china_fit, aes(x=value/365.25*12, y=event, color=event_type, fill=event_type), alpha=0.3) +
  geom_point(data=indv_2, aes(x=marginal_delay/365.25*12, y=event, color=event_type, group=Volunteer)) +
  scale_x_continuous(breaks = 0:15, limits=c(0, 15)) +
  scale_color_discrete(drop=F) +
  scale_fill_discrete(drop=F) +
  labs(title = "Individual volunteer infection delays",
       subtitle = paste("Densities show fit with", names(samples_list)[5]),
       x = "Months since previous event",
       y = NULL,
       color = "Event type",
       fill = "Event type")

ggsave("../plots/volunteer_relapses.png", width=8, height=4)
```

How about the proportion of long relapses?

```{r}
short_incubation_people = indv %>%
  filter(event_num == 0,
         !is_long)

long_relapses = indv %>%
  filter(Volunteer %in% short_incubation_people$Volunteer) %>%
  filter(event_num == 1,
         is_long)

short_relapses = indv %>%
  filter(Volunteer %in% short_incubation_people$Volunteer) %>%
  filter(event_num == 1,
         is_short)
proportion = nrow(long_relapses) / nrow(short_relapses)

p_long = seq(0, 1, length.out=100)
yy = dbeta(p_long, nrow(long_relapses), nrow(short_relapses))
plot(p_long, yy, 'l')
```
