---
title: "Scrape tutiempo.net"
output: html_notebook
---

This RMarkdown notebook is for scraping weather data from en.tutiempo.net.

Note: consider moving to this dataset instead:
https://crudata.uea.ac.uk/cru/data/hrg/?_ga=2.86119323.20970249.1688623809-2016247340.1688623809

```{r setup}
library(tidyverse)
library(RSelenium)
library(rvest)
library(lubridate)
library(pbapply)
library(memoise)
library(MalariaData)
R_files = list.files("../R/functions", "\\.R$", recursive=TRUE, full.names=TRUE)
for (file in R_files) { source(file) }

selection = "Shangqiu"

load_functions = list(Guantang = function(...) load_region("guantang", ...),
                      Yingjiang = function(...) load_region("yingjiang", ...),
                      Dengzhou = function(...) load_region("deng", ...),
                      Huangchuan = function(...) load_region("huangchuan", ...),
                      Shangqiu = function(...) load_region("shangqiu", ...),
                      Tengchong = function(...) load_region("tengchong", ...))
URIs = list(Tengchong = "ws-567390.html",
            Shangqiu = "ws-580050.html")

base_url = "https://en.tutiempo.net/climate/"
date_fmt = "%m-%Y"

form_url = function(date, base_url="https://en.tutiempo.net/climate", uri="ws-567390.html") {
  file.path(base_url, date, uri)
}
epi_data = load_functions[[selection]]()
dates = seq(min(epi_data$Date), max(epi_data$Date), by="month") %>%
  format(date_fmt) %>%
  setNames({.})
URLs = sapply(dates, form_url, uri=URIs[[selection]])
```

```{r}
.get_table_data = function(url) {
  html = read_html(url)
  table_data = html %>%
    html_node("table.medias") %>%
    html_table() %>%
    filter(as.character(suppressWarnings(as.numeric(Day))) == Day) %>% # ensure row has a real Day, not a gap or total
    mutate_all(\(...) suppressWarnings(as.numeric(...)))
}
mem = cache_filesystem("cache")
get_table_data = memoise(.get_table_data, cache=mem)
try_get_table_data = function(...) {
  try(get_table_data(...))
}
```

```{r}
tables = pblapply(URLs, try_get_table_data)
table_success = sapply(names(tables), function(x) {
  ifelse("try-error" %in% class(tables[[x]]), NA, x)
}) %>% unname()
table_success = table_success[!is.na(table_success)]
tables_filt = tables[table_success]
```

```{r}
weather = bind_rows(tables_filt, .id="Month") %>%
  mutate(Date = paste0(Day, "-", Month) %>% dmy()) %>%
  select(Date, everything(), -Month, -Day) %>%
  mutate(DayInYear = Date - ymd(paste0(year(Date), "01-01")) + 1)


seasonal = weather %>%
  drop_na(`T`, TM, Tm) %>%
  group_by(DayInYear) %>%
  # summarise(across(`T`:VM, \(x) mean(x, na.rm=T)))
  summarise(average_mean = mean(`T`, na.rm=T),
            average_max = mean(`TM`, na.rm=T),
            average_min = mean(`Tm`, na.rm=T),
            average_precip = mean(PP, na.rm=T),
            TEST = mean(`T`, na.rm=T))

weather %>%
  rename(Average = `T`,
         Max = TM,
         Min = Tm) %>%
  pivot_longer(cols=c(Average, Max, Min)) %>%
  drop_na(value) %>%
  ggplot(aes(x=ymd("1999-12-31")+DayInYear, y=value, color=name, group=interaction(name, year(Date)))) +
  geom_line() +
  scale_x_date(date_labels="%b", date_breaks="month") +
  labs(title=paste(selection, min(weather$Date), "to", max(weather$Date)), x = "Month", y="deg Celsius")

seasonal %>%
  ggplot(aes(x=ymd("1999-12-31")+DayInYear, y=average_mean)) +
  geom_point() +
  geom_point(aes(y=average_precip*max(seasonal$average_mean,na.rm=T)/max(seasonal$average_precip, na.rm=T)), color="steelblue") +
  scale_x_date(date_labels="%b", date_breaks="month") +
  scale_y_continuous(sec.axis = sec_axis(~. *max(seasonal$average_precip, na.rm=T)/max(seasonal$average_mean, na.rm=T),
                                         name = "Precipitation (mm)")) +
  theme(axis.ticks.y.right = element_line(color = "steelblue"),
        axis.text.y.right = element_text(color = "steelblue"),
        axis.title.y.right = element_text(color = "steelblue")) +
  labs(title=paste(selection, min(weather$Date), "to", max(weather$Date)), x = "Month", y="Average daily mean temperature (C)")
ggsave(paste0("plots/", selection, "_climate.svg"), width=6, height=4)
```

Variable | Description
---------|------------
       T | Average Temperature (°C)
      TM | Maximum temperature (°C)
      Tm | Minimum temperature (°C)
     SLP | Atmospheric pressure at sea level (hPa)
       H | Average relative humidity (%)
      PP | Total rainfall and / or snowmelt (mm)
      VV | Average visibility (Km)
       V | Average wind speed (Km/h)
      VM | Maximum sustained wind speed (Km/h)
      VG | Maximum speed of wind (Km/h)
      RA | Indicate if there was rain or drizzle (In the monthly average, total days it rained)
      SN | Snow indicator (In the monthly average, total days that snowed)
      TS | Indicates whether there storm (In the monthly average, Total days with thunderstorm)
      FG | Indicates whether there was fog (In the monthly average, Total days with fog)
